---
title: "datainsightR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datainsightR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datainsightR)
library(dplyr)
```

***Introduction

The datainsightR package provides lightweight tools for quick data cleaning
and exploratory analysis. It is designed to help users summarize missing values,
clean column names, detect outliers, and generate basic descriptive statistics
with minimal code.

This vignette demonstrates the core functions and how they can be used
in a typical workflow.


**1. Summarizing Missing Data

The summarize_missing() function counts the number of missing values
in each column of a data frame.

```{r}
df <- data.frame(
a = c(1, 2, NA, 4),
b = c("x", NA, "y", "z")
)

summarize_missing(df)
```

**2. Cleaning Column Names

The clean_names() function standardizes column names by:

converting to lowercase

replacing special characters with _

removing leading/trailing underscores
```{r}
messy <- data.frame(
"First Name" = 1:3,
"Value-Amount$" = c(10, 20, 30)
)

clean_names(messy)
```

**3. Detecting Outliers

The debug_outliers() function identifies numeric outliers
based on the 1.5 × IQR rule.
```{r}
values <- c(1, 2, 3, 4, 100, 150)
debug_outliers(values)
```

**4. Quick Descriptive Statistics

The quick_describe() function computes:

mean

median

standard deviation

minimum

maximum

for all numeric columns.
```{r}
quick_describe(mtcars)
```

***Conclusion

The datainsightR package provides a compact set of utilities
that streamline basic data cleaning and exploratory analysis tasks.

Future extensions may include:

correlation heatmaps

automated EDA reports

more robust data validation tools

Thank you for using datainsightR!
